---
title: "Graph examples with penguins"
author: "James Glover"
date: "20/08/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(gt)
library(ggthemes)

theme_set(theme_light())
#theme_set(theme_dark())

knitr::opts_chunk$set(echo = FALSE)
```

### Get the data

# These examples use the simplified 'penguins_size.csv' availabe in the repo. 

```{r warning=FALSE}
penguins <- read_csv(here("data","penguins_size.csv"))
```
## Example graphs

#### Penguin mass vs. flipper length:

```{r warning=FALSE}
ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 2,
             alpha = 0.9) +
  #theme_void() +
  theme_minimal() +
  #theme_few() +
  #theme_calc() + 
  #theme_classic() +
  #scale_color_manual(values = c("#A8AD00","#6699CC","#C5003E")) +
  #scale_fill_brewer(palette="Set1") +
  #scale_color_brewer(palette = "Blues") + 
  scale_color_brewer(type	= 'qual', palette = 'Dark2') + 
  labs(title = "Flipper length and body mass",
       subtitle = "The Gentoo Penguins have the largest Flippers and Body Mass",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.2, 0.75),
        legend.background = element_rect(fill = "white", color = NA),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

```

#### Flipper length vs. culmen length

```{r warning=FALSE}
ggplot(data = penguins,
                         aes(x = body_mass_g,
                             y = culmen_length_mm)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 2,
             alpha = 0.9) +
  theme_minimal() +
  scale_color_brewer(type	= 'qual', palette = 'Dark2') + 
  labs(title = "Flipper and Body Mass",
       subtitle = "The Chinstrap Pengiuns generally have larger Culmens (Beaks)",
       x = "Flipper length (mm)",
       y = "Culmen length (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.88, 0.2),
        legend.background = element_rect(fill = "white", color = NA),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

```
#### Flipper length vs. culmen length

```{r warning=FALSE}
ggplot(data = penguins,
       aes(x = culmen_length_mm
           ,y = culmen_depth_mm
           ,group = species
           )) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 2,
             alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, aes(color = species)) +
  theme_light() + 
  scale_color_brewer(type	= 'qual', palette = 'Dark2') + 
  labs(title = "Culmen (Breaks) length and depth",
       subtitle = "Each Groups breaks depth increases with length",
       x = "Culmen length (mm)",
       y = "Culmen depth (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.9, 0.17),
        legend.background = element_rect(fill = "white", color = NA),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

```

#### Simpson's paradox, omitting species in graph above:  

```{r warning=FALSE}
ggplot(data = penguins,
       aes(x = culmen_length_mm
           ,y = culmen_depth_mm
           )) +
  geom_point(colour = "cyan4", size = 1, alpha = 1, shape = 1) +
  theme_light() + 
  labs(title = "Simpson's paradox",
       subtitle = "However the data suggests that the depth decreases with length",
       x = "Culmen length (mm)",
       y = "Culmen depth (mm)") +
  geom_smooth(method = "lm", se = FALSE, color = "cyan4") 

```

#### Some other graphs for brainstorming

Flipper lengths histogram:  

```{r warning=FALSE}

ggplot(data = penguins, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species), 
                 alpha = 0.65, 
                 position = "identity") +
  scale_fill_brewer(palette = 'Dark2') + 
  theme_minimal() +
  labs(x = "Flipper length (mm)",
       y = "Frequency",
       title = "Penguin flipper lengths",
       subtitle = "Gentoo have the largest flippers"
       )

```

Flipper lengths density plot:  

```{r warning=FALSE}

ggplot(data = penguins, aes(flipper_length_mm, fill = species)) +
  geom_density(alpha = 0.65) +
  scale_fill_brewer(palette = 'Dark2') + 
  theme_light() + 
  #theme_minimal() +
  labs(x = "Flipper length (mm)",
       y = "Frequency",
       title = "Penguin flipper lengths",
       subtitle = "Gentoo have the largest flippers")

ggplot(data = penguins, aes(flipper_length_mm,after_stat(count), fill = species)) +
  geom_density(position = "fill", alpha = 0.5) +
  scale_fill_brewer(palette = 'Dark2') + 
  theme_light()

ggplot(data = penguins, aes(flipper_length_mm,after_stat(count), fill = species)) +
  geom_density(position = "stack", alpha = 0.5) +
  scale_fill_brewer(palette = 'Dark2') + 
  theme_light()

```

Flipper lengths boxplots + jitter: 

```{r warning=FALSE}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(aes(color = species), width = 0.3, show.legend = FALSE) +
  geom_jitter(aes(color = species), width = 0.1, alpha = 0.65, show.legend = FALSE) +
  scale_color_brewer(palette = 'Dark2') + 
  theme_light() + 
  labs(x = "Species",
       y = "Flipper length (mm)",
       title = "Penguin flipper lengths",
       subtitle = "Gentoo have the largest flippers"
       )


ggplot(data = penguins, aes(species, flipper_length_mm)) +
    geom_boxplot(aes(color = species), width = 0.3, show.legend = TRUE) +
    geom_jitter(aes(color = species), width = 0.1, size=0.4, alpha = 0.65, show.legend = FALSE) +
    scale_color_brewer(palette = 'Dark2') + 
    theme_light() + 
  labs(y = "Flipper length (mm)",
       title = "Penguin flipper lengths",
       subtitle = "Gentoo have the largest flippers"
       ) + 
  theme(
    axis.text.x = element_blank(),
    legend.position = c(0.82, 0.2),
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 0, face= "italic"),
    plot.caption.position = "plot")

```

Body mass histograms: 

```{r warning=FALSE}

ggplot(data = penguins, aes(x = body_mass_g)) +
  geom_histogram(aes(fill = species), 
                 alpha = 0.5, 
                 position = "identity") +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  theme_light() + 
  labs(x = "Body Mass (g)",
       y = '',
       title = "Penguin Body Mass",
       subtitle = "Gentoo are the largest Penguin"
       ) + 
  theme(
    axis.text.x = element_blank(),
    legend.position = c(0.82, 0.8),
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 0, face= "italic"),
    plot.caption.position = "plot")

```

### Include penguin sex: 
```{r warning=FALSE}
penguins %>% 
  filter(sex %in% c("FEMALE", "MALE")) %>% 
ggplot(., aes(x = flipper_length_mm,
                            y = body_mass_g)) +
  geom_point(aes(color = sex), alpha = 0.6) +
  theme_light() + 
  scale_color_manual(values = c("red","blue")) +
  labs(title = "Penguin flipper and body mass",
       subtitle = "The Males appear to be the larger of the two groups",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Sex") +
  theme(legend.position = c(0.85, 0.18),
        legend.background = element_rect(fill = "white", color = NA),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot"
        ) +
  facet_wrap(~species)
```


```{r warning=FALSE}
penguins %>% 
  filter(sex %in% c("FEMALE", "MALE")) %>% 
  ggplot(., aes(x = flipper_length_mm,
                            y = body_mass_g)) +
  geom_boxplot(aes(fill=sex)) + 
  theme_minimal()

penguins %>% 
  filter(sex %in% c("FEMALE", "MALE")) %>% 
  
  ggplot(., aes(x = flipper_length_mm,
                            y = body_mass_g)) +
  geom_boxplot(aes(color = sex), alpha = 0.6) +
  geom_jitter(aes(color = sex), alpha = 0.6) + 
  theme_light() + 
  scale_color_manual(values = c("red","blue")) + 
  labs(x = "Flipper length (mm)",
       y = "Body mass (g)",
       title = "Penguin flipper and body mass",
       subtitle = "The Males appear to be the larger of the two groups"
       ) + 
  theme(
    axis.text.x = element_blank(),
    legend.position = c(0.82, 0.2),
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 0, face= "italic"),
    plot.caption.position = "plot"
 ) +
  facet_wrap(~species)
```
